---
title: "Midpoint Deliverable"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("knitr")
# source("scripts/summary_info.R")
# source("scripts/summary_table.R")
# source("scripts/chart1.R")
# source("scripts/chart2.R")
source("scripts/chart3.R")
bac <- read.csv("data/dataset1.csv")
colnames(bac)[colnames(bac)=="Driver.BAC...0.00..Percent"] <- "Percent_Drivers_0_BAC"
colnames(bac)[colnames(bac)=="Driver.BAC....08...Percent"] <- "Percent_Drivers_over_0.08_BAC"
library("ggplot2")
library("plotly")
```

Leverage the functions written in the other files to create your report:

## Alcohol and car accidents

### Introduction

According to National Public Radio ([NPR](https://www.npr.org/sections/thetwo-way/2017/08/10/542409957/drinking-on-
the-rise-in-u-s-especially-for-women-minorities-older-adults)), alcohol consumpt
ion has been increasing in the United States, and out group wants to explore how 
this change may be related to the frequency of car accidents. 
One of the data sets we are using to examine this issue is "Drivers Involved in 
Fatal Crashes, by State and Blood Alcohol Concentration of the Driver" ([source](https://www-fars.nhtsa.dot.gov/States/StatesAlcohol.aspx)). This 
data covers the United States, up to the year 2017, and was collected by the 
Fatality Analysis Reporting System.

* Write an introductory paragraph explaining the domain/data set (feel free to borrow from the brainstorm part of the assignment)
* Write a summary paragraph that includes information calculated by your summary information function
* Render your table and describe pertinent observations found in it
* Describe the purpose of each chart (e.g., "this chart was intended to show the relationship between..."), render each chart, and describe pertinent observations found in each chart
* In order to render this information, you'll need to read in your data in (only) your index.Rmd file, and pass your data to the functions provided in each script.

## R Markdown

This third chart was intended to show the relationship between drivers involved in fatal crashes with zero blood alcohol content compared to those who were over the legal limit of 0.08 BAC. This makes an effort to answer, _"Is a driver with higher blood alcohol concentration (BAC) more likely to be involved in a crash?"_ 
Each point represents a state.

```{r chart3, echo=FALSE}
chart3(bac)
```

Here, we see that in most states, over 70% of drivers involved in fatal crashes were 
sober (had blood alcohol content of 0), shown on the right half of the plot. There are a few exceptions, shown on the left half of the plot, where there is a lower percentage of sober drivers in accidents and a higher number of drivers over the legal limit.