---
title: "Midpoint Deliverable"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# load libraries
library("knitr")
library("choroplethr")
library("choroplethrMaps")
library("ggplot2")
library("plotly")

# load source files
source("scripts/summarize_info.R")
source("scripts/make_table.R")
source("scripts/Histogram_Fatalities.R")
source("scripts/chart2.R")
source("scripts/chart3.R")

# load data
death_rate <- read.csv(
  "data/Impaired_Driving_Death_Rate__by_Age_and_Gender__2012___2014__All_
  States.csv",
  stringsAsFactors = FALSE)
bac <- read.csv("data/dataset1.csv", stringsAsFactors = FALSE)
ny_3yr_crashes <- read.csv("data/mvc_threeyear_ny.csv")

# getting summary info from ny crashes
ny_crashes <- get_summary_info(ny_3yr_crashes)

# rename columns
colnames(bac)[1] <- "State"
colnames(bac)[colnames(bac) == "Driver.BAC...0.00..Percent"] <-
  "Percent_Drivers_0_BAC"
colnames(bac)[colnames(bac) == "Driver.BAC....08...Percent"] <-
  "Percent_Drivers_over_0.08_BAC"
bac <- head(bac, -1) # remove the "total" row

# creating table
bac_table <- table_maker(bac)
```

## Alcohol and car accidents

### Introduction

According to National Public Radio ([NPR](https://www.npr.org/sections/thetwo-way/2017/08/10/542409957/drinking-on-
the-rise-in-u-s-especially-for-women-minorities-older-adults)), alcohol consumption has been increasing in the United States, and our group wants to explore how 
this change may be related to the frequency of car accidents. 
One of the data sets we are using to examine this issue is "Drivers Involved in 
Fatal Crashes, by State and Blood Alcohol Concentration of the Driver" ([source](https://www-fars.nhtsa.dot.gov/States/StatesAlcohol.aspx)). This 
data covers the United States, up to the year 2017, and was collected by the 
Fatality Analysis Reporting System.

### Summary Paragraph
For our datasets, the datasets we used gave every car accident for a certain state/region for a given year. For our chosen set, it reported every accident in New York from 2014 to 2016. The summary information that we calculated was the number of accidents per year, the number of accidents with alcohol has a cause, the percentage of accidents caused by alcohol, average accidents per year, and most common day car type. From 2014 to 2016, New York had `r ny_crashes$total_crashes` crashes and `r ny_crashes$total_alc_crashes` were caused by drivers under the influence of alcohol. This is a large number of car accidents due to alcohol consumption, but it was only `r ny_crashes$pct_crash`% of the total crashes in New York. While it is only a small percentage of New York's total crashes, the number itself is significant and can be reduced when people cease drunk driving. On average, New York had `r as.integer(ny_crashes$avg)` crashes per year, and the year with the highest crashes was `r ny_crashes$highest_year`, showing that the number of crashes has not been decreasing.

### Summary Table
This table shows the total drivers involved in fatal car accidents up to 2017 and the percentage of drivers that had a Blood Alcohol Content (BAC) greater than or equal to 0.1 by state. The table is arranged by decreasing number of fatal deaths. This information is displayed to show how many fatal crashes can be reduced if drivers are responsible and do not drunk drive.

``` {r}
kable(bac_table)
```

From the table, one can see that a high percentage of states' fatal car crashes are due to drivers with a BAC greater than or equal to 0.1. The top three states by number of fatal crashes are Texas, California, and Florida. But the table also shows that the state size (population) correlates to the number of accidents. The most important takeaway is knowing that roughly `r round(mean(bac_table$"Percent of Drivers w/ BAC over 0.1"), digits = 2)`% of fatal car crashes in all states is due to drunk driving, meaning that these crashes are all preventable if drivers act responsibly. 

### Charts
The bar plot below displays the percentage of adults involved in car accident fatalities in the United States sorted by BAC (Blood Alcohol Concentration). The three bars have information at the 0, 0.1-0.7, 0.8+ BAC levels respectively. 
```{r chart1, echo = FALSE, warnings = FALSE}
n_data(bac)
```

As can be seen, this chart clearly depicts how the the highest frequencies in adult fatalities were at the 0.0 BAC level and 0.8+ level. This shows that there were more US adults involved in these car accidents when they were legally impaired(0.8+), as compared to intoxicated at a BAC level ranging from 0.1-0.7. Clearly, the most extreme BAC levels have led to some of the highest percentage of adults involved in these car accident fatalities.

The next chart includes multiple graphs of fatal car accident percentage by age and gender accross the United States. This chart intend to show _"The relationship between driving under the influence and age/gender in 2012 and 2014."_
```{r chart2, echo=FALSE, message=FALSE, warning=FALSE, results="hide"}
# 2012/2014, 0-20/21-34/35+/male/female
chart2(death_rate, 2014, "male")
```

From these graphs, we can see that female drivers are less likely to be involved 
in fatal drunk driving accidents across the United States, while drivers between 
the ages of 21 and 35 are most likely to be involved in fatal drunk driving accidents. Overall, states like MT, ND, WY, and NM have higher drunk driving 
death rates, and states like WA, MI and UT have lower death rates. This general
trend hasn't changed from 2012 to 2014. 

This third chart was intended to show the relationship between drivers involved in fatal crashes with zero blood alcohol content compared to those who were over the legal limit of 0.08 BAC. This makes an effort to answer, _"Is a driver with higher blood alcohol concentration (BAC) more likely to be involved in a crash?"_ 
Each point represents a state.

```{r chart3, echo = FALSE, warnings = FALSE}
chart3(bac)
```

Here, we see that in most states, over 70% of drivers involved in fatal crashes were sober (had blood alcohol content of 0), shown on the right half of the plot. There are a few exceptions, shown on the left half of the plot, where there is a lower percentage of sober drivers in accidents and a higher number of drivers over the legal limit.